---
title: "R Markdown and Leaflet Submit project"
author: "Abhinav Agarwal"
date: "8 December 2018"
output: html_document
---


# Introduction
This is a Coursera project in Developing Data Products course
Create a web page using R Markdown that features a map created with Leaflet.

Host your webpage on either GitHub Pages, RPubs, or NeoCities.

Your webpage must contain the date that you created the document, and it must contain a map created with Leaflet. We would love to see you show off your creativity!

# Data plotted


The data used comes from https://simplemaps.com/data/world-cities

This data shows about 13,000 cities around the world. 

```{r}
data<-read.csv("worldcities.csv", header=TRUE, sep=",")

library(leaflet)
```

```{r}
my_map <- data %>%
        leaflet() %>%
        addTiles() %>%
        setView(lng = 0, lat = 0,zoom=1) %>%

        addCircleMarkers(popup = data$city, 
                   lng = data$lng, 
                   lat = data$lat,radius = 0.05) 
        

my_map
```


